---
title: "Quality Assurance and Control: Pt. 2 Calibration"
output:
  html_document:
    df_print: paged
---

# TODO: Chunk models into time-periods.
- Pre/Post Covid 19. 
- Season.

```{r environment setup, include=FALSE}
library(dplyr)
devtools::load_all()
library("AirSensor")
library(ropenaq)
library(ggplot2)
source("C://Users/iozeroff/Data-Science/R-Projects/useful/model-equation.R")
source("C://Users/iozeroff/Data-Science/R-Projects/useful/model-mutate-functions.R")
library(modelr)
library(ggthemr)
library(stringr)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(purrr)

knitr::opts_knit$set(root.dir = here::here())

# Establish Plotting Theme
ggthemr::ggthemr("pale", layout = "clean", type = "outer")
```

```{r, include=FALSE}
site <- "IMD Lodhi Road"

sensor_catalog <- load_SensorCatalog(site = site)
pat_list <- load_pat_list(site = site)
source("scripts/qaqc.R")

# Get Collocation Site Name from Sensor Catalog
ref_site <- sensor_catalog %>% 
  distinct(`Collocation Site`) %>%
  pull()
```


As described in Part 1 of this vignette, Operation Healthy Air utilizes a rigid cleaning process to remove data points that are invalid or unlikely. This allows greater confidence in the trends being displayed by one or more our sensors. Purple Air sensors are low-cost, and have vulnerabilities more expensive sensors, such as Federal Reference Monitors deal with, most imporantly different sensitivities to different particle shapes.

Briefly, Plantower Sensors (the PM detectors used in Purple Air Sensors) shine lasers through air samples and 'count' the Particulate Matter 2.5 within a sample. A key aspect of how this 'count' is taken, is assuming all particulates are spheres. Not all particulates are spheres. 

As a result, taken by Purple Air sensors vary dependent on the pollution source and type. Wood burning smoke and vehicle exhaust emissions in the same density can register different values by a PM sensor.

Our solution is to calibrate sensors to local conditions using Federal Reference Monitors (cite). Operation Healthy Air programs in Boston and India do not yet have the resources to calibrate each sensor before deployment. To compensate, we calibrate by proxy.Sensors at Federal Reference Monitor Sites in Boston and Delhi create calibrations using these sensors. We apply our program calibration curves to program sensors, assuming conditions are relatively similar within a programs geography. 

## Reference Monitor Data

Here we'll calibrate our senses at tje `r print(ref_site)` site.

We import Reference Monitor Data from the OpenAQ API (https://docs.openaq.org/). Currently, the API only provides the previous 90 days of data. We are working to access more reference monitor data to create more robust calibration curves.

```{r, include = FALSE}
# Site metadata.
ref_meta <- aq_locations() %>% filter(location == ref_site)

# Get Collocation Site Reference Monitor Data  
ref_data <- aq_measurements(location = ref_meta$locationURL, parameter = "pm25") %>% 
  select(datetime = "dateUTC", pm25 = value) %>% 
  filter(pm25 >= 0)
```


Our calibration curves are multivariate linear models--we create an equation that accounts for sensor temperature and humidity to transform our raw PM measurement towards the same scale as a Federal Reference Monitor.

```{r}
# Loading AirSensor objects.
sensor_list <- readRDS("data/sensor_list.rds")

# Subset AirSensors with PM data.
pm_data <- sensor_list %>% 
  map("pm25") %>% 
  map(.f = function(x) dplyr::rename(x$data, pm25 = 2))
# Subset AirSensors with Temp data.  
temp_data <- sensor_list %>% 
  map("temperature") %>% 
  map(.f = function(x) rename(x$data, temperature = 2))
# Subset AirSensors with Humidity data.
humidity_data <- sensor_list %>% 
  map("humidity") %>% 
  map(.f = function(x) rename(x$data, humidity = 2))

# Join PM Data to Reference Data
pm_data <- pm_data %>% 
  map(.f = function(x) inner_join(x = x,
                                  y = ref_data,
                                  by = "datetime",
                                  suffix = c("_pa", "_ref"))
  )

# Combining all sensor variables into single dataframes for modelling.
data <- 
  map2(.x = pm_data, .y = temp_data,
       .f = left_join, by = "datetime") %>%
  map2(.y = humidity_data, 
       .f = left_join, by ="datetime") %>%
  # Filter complete cases.
  map(.f = function(x) x[complete.cases(x), ])

```


We have `r length(sensor_list)` at the site, so we generate curves for each. 

```{r}
# Make Models
test2 <- data %>% 
  map(
      .f = function(x) lm(pm25_ref ~ pm25_pa + temperature + humidity, data = x)
    )

map_df(.x = models, .f = function(x) model_equation(x)) %>% 
  knitr::kable()
```

Collocating multiple sensors, allows us to select the robust model from among multiple options. But, model selection is no simple task. 

Below are several model selection criteria that may be used to select a calibration curve. Operation Healthy Air uses the adjusted R-squared statistic to select the best calibration curve. 

```{r model summary stats, message=FALSE, echo=FALSE}
# TODO: Add Root Mean Square Error
mod_summaries <- 
  data.frame(
    label = names(models),
    p_value = map_dbl(.x = models, .f = extract_p),
    AIC = map_dbl(.x = models, .f = AIC),
    R2 = map_dbl(.x = models, .f = extract_R2, adjusted = FALSE),
    R2_adjusted = map_dbl(.x = models, .f = extract_R2, adjusted = TRUE),
    RMSE = map_dbl(.x = models,
                    ~ sqrt(mean((.x$residuals) ^ 2))
                   ),
    missing = map_dbl(.x = pm_data, .f = function(x) sum(is.na(x$pm25_pa))),
    row.names = NULL
    ) 

mod_summaries %>%
  knitr::kable()
```


```{r Model Correction Visualization, message=FALSE, echo=FALSE}
# Get label of best model.
proxy_sensor <- mod_summaries %>% 
  filter(R2 == (max(R2))) %>%
  pull(label)
# Select best model.
proxy_model <- models[[proxy_sensor]]
# Visualize Transformation
broom::augment(proxy_model) %>%
   select(position = .rownames, pm25_pa, pm25_ref, .fitted) %>%  
  mutate(
    position = as.integer(position),
  ) %>%
  tidyr::pivot_longer(-position) %>%
  mutate(
    name = factor(
      name,
      levels = c(".fitted",
                 "pm25_pa",
                 "pm25_ref"),
      labels = c("Calibrated",
                 "Collocated",
                 "Reference"),
    )
  ) %>% 
  ggplot(aes(x = position, y = value)) +
  geom_point(aes(color = name), position = "jitter", alpha = 0.5) +
  geom_line(aes(color = name), position = "jitter", alpha = 0.8, size = 1.2) +
  labs(
    fill = "Data",
    color = "Data",
    y = "µg/m3",
    title = "Calibration",
    caption = paste("FRM:", ref_site),
    subtitle = proxy_sensor
  ) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "top",
        legend.title = element_blank()
        )
```

## Application

Once we've selected a model we feel confident in, we apply it to our cleaned and calibrated sensor data.

```{r}
# Load another sites raw data.
pat_list <- load_pat_list(site = "Modern Public School")

# QAQC
suppressMessages({
  source("scripts/qaqc.R")
})

```

Cleaned and calibrated data comes as 2-minute increment time series of channel A and Channel B measurements. We calibrate each of these channels, and the aggregate data into an hourly dataset. Our high resolution dataset can also be used to observe hyperlocalized events. 

```{r}

pat_list <- readRDS("data/pat_list_qcd.rds")

pa_meta <- pat_list %>% 
  map(.f = pat_extractMeta) %>% 
  bind_rows(.id = "label")

ref_map <- leaflet() %>%
  addTiles() %>%
  addMarkers(
    data = ref_meta, 
    icon = list(
      iconUrl = 'http://icons.iconarchive.com/icons/artua/star-wars/128/Master-Joda-icon.png',
      iconSize = c(75, 75)
    ),
    lat  = ~latitude, lng = ~longitude,
    label = paste("Reference Monitor",
                  ~htmlEscape(location), sep = "\n")
    ) %>% 
  addCircleMarkers(data = pa_meta, color = "purple",
                   lat = ~latitude, lng = ~longitude, 
                   label = ~htmlEscape(label))

print(ref_map)

pat_list_calibrated <- list()

for (i in 1:length(pat_list) ) {
  
  # To calibrate a PAT, must calibrate each channel individually.
  pat <- pat_list[[i]]
  
  meta <- pat %>%
    pat_extractMeta()
  
  data <- pat %>%
    pat_extractData() 
  
  data[["pred_A"]] <- data %>% 
    select(pm25_pa = pm25_A, temperature, humidity) %>% 
    modelr::add_predictions(model = proxy_model) %>% 
    pull(pred)
    
  data[["pred_B"]] <- data %>% 
    select(pm25_pa = pm25_B, temperature, humidity) %>% 
    modelr::add_predictions(model = proxy_model) %>% 
    pull(pred)
  
  
  plot <- data %>% 
    select(datetime, pm25_A, pm25_B, pred_A, pred_B) %>%  
    tidyr::pivot_longer(-datetime) %>% 
    mutate(
      channel = if_else(str_detect(name, "_A$"), "A", "B"),
      calibrated = str_remove(name, "_[AB]$"),
      calibrated = factor(
        calibrated,
        levels = c("pred",
                   "pm25"),
        labels = c("Calibrated",
                   "Uncalibrated")
      )
    ) %>% 
    ggplot(aes(x = datetime,
               y = value,
               color = calibrated)) +
    geom_line(position = "jitter",
              alpha = 0.5, size = 1.2) + 
    labs(
      color = "Calibrated",
      y = "µg/m3",
      title = "Calibration",
      caption = paste("FRM:", ref_site),
      subtitle = meta$label
    ) +
    facet_grid(channel~.) +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      legend.position = "top",
      legend.title = element_blank(),
      legend.background = element_blank()
      ) +
    guides(
      color = guide_legend(override.aes = list(size = 5))
      )
  
  print(plot)
  
  
  pat[["data"]] <- data %>% 
    select(-pm25_A, -pm25_B) %>% 
    rename(pm25_A = pred_A,
           pm25_B = pred_B)    
  
  pat_list_calibrated[[meta$label]] <- pat
    
}

```

And we can see perhaps more clearly how our hourly data is affected by calibration as well.

```{r echo=FALSE, message=FALSE}

uncalibrated <- readRDS("data/sensor_list.RDS") %>%
  map(.f = function(x) x$pm25$data) %>% 
  reduce(.f = full_join, by = "datetime") %>% 
  tidyr::pivot_longer(-datetime, names_to = "sensor")

calibrated <- map(pat_list_calibrated,
                 .f = function(x) pat_createAirSensor(x)$data) %>% 
  reduce(.f = full_join, by = "datetime") %>% 
  tidyr::pivot_longer(-datetime, names_to = "sensor")

bind_rows(uncalibrated, calibrated, .id = "calibrated") %>%
  mutate(
    calibrated = factor(
      calibrated, 
      levels = c(1, 2),
      labels = c("Uncalibrated", "Calibrated"))
  ) %>%
  ggplot(aes(x = datetime,
             y = value,
             color = calibrated)) +
  geom_line(position = "jitter",
              alpha = 0.5, size = 1.2) + 
  labs(
    color = "Calibrated",
    y = "µg/m3",
    title = "Calibration",
    caption = paste("FRM:", ref_site),
    subtitle = meta$label
  ) +
  facet_grid(sensor~.) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    legend.position = "top",
    legend.title = element_blank(),
    legend.background = element_blank()
    ) +
  guides(
    color = guide_legend(override.aes = list(size = 5))
    )

```

